Pretty darn good control: when are approximate solutions better than approximate models

### 2023-06-14 manuscript/manuscript.Rmd
changes in math model

In this section we introduce some models describing the population dynamics of fisheries.
In general, the class of models that appear in this context are *first order finite difference equations* (potentially stochastic difference equations).
For $n$ species, these models have the general form
\begin{align}
  \label{eq:general model}
  \begin{split}
    \Delta X_t &= f_X(N_t) - M_X X_t\\
    \Delta Y_t &= f_Y(N_t) - M_Y Y_t\\
    &\dots
  \end{split}
\end{align}
where $N_t = (X_t,\ Y_t,\ \dots) \in \mathbb{R}^{n}_+$ is a vector of populations, $\Delta X_t := X_{t+1} - X_t$, $\Delta Y_t := Y_{t+1} - Y_t$, and where $f_i:\mathbb{R}^{n}\to\mathbb{R}$ are arbitrary functions.
Here, $M_i=M_i(N_t)\in\mathbb{R}_+$ is a state-dependent fish mortality arising from harvesting (sometimes referred to as *fishing effort*).
Thus, the term $M_X X_t$ is the total $X$ harvest at time $t$.
We will focus on systems where a single species, $X$, is harvested, i.e. $M_Y = M_Z = \dots = 0$.

@@

```{r}
#| label: may
#| fig.cap: "The fixed point diagram for the model associated to $f_{\\text{May}}$ as a function of varying the parameter $\\beta H$.
#| That is, these are the zeroes of the function $f_{\\text{May}}$. 
#| Stable fixed points (also known as *attractors*) are plotted using a solid line, while the unstable fixed point is shown as a dotted line."
#| out.width: "6.5in"

@@

Consider the following three species generalization of eq. \eqref{eq:may}:

\begin{align}
\label{eq:3d model}
\begin{split}
  \Delta X_t &= L(X_t;\ r_X, K_X) - F(X_t, Z_t;\ \beta, c) - c_{XY}X_tY_t - h_t,\\
  \Delta Y_t &= L(Y_t;\ r_Y, K_Y) - D F(Y_t, Z_t;\ \beta, c) - c_{XY}X_tY_t,\\
  \Delta Z_t &= \left(f(X_t + DY_t\right) - d_Z)Z_t.
\end{split}
\end{align}
The model is conceptualized in Fig. \ref{fig:conceptual}, panel B.
It contains three populations, the state at time $t$ is $(X_t, Y_t, Z_t)$. 
Species $Z$ preys on both $X$ and $Y$, while the latter two compete for resources.
There are ten parameters in this model: 
The growth rate and carrying capacity, $r_X$, $K_X$, $r_Y$ and $K_Y$, of $X$ and $Y$.
A parameter $c_{XY}$ mediating a Lotka-Volterra competition between $X$ and $Y$.
A maximum predation rate $\beta$ and a predation rate half-maximum biomass $c$ specifying how $Z$ forages on $X$ and $Y$.
A parameter $D$ regulating a relative preference of $Z$ to prey on $Y$.
Finally, a death rate $d_Z$ and a parameter $f$ scaling the birth rate of $Z$.

>>>>

In this section we introduce some models describing the population dynamics of fisheries.
In general, the class of models that appear in this context are *first order finite difference equations* (potentially stochastic difference equations).
For $n$ species, these models have the general form
\begin{align}
  \label{eq:general model}
  \begin{split}
    \Delta X_t &= f_X(N_t) - M_X X_t\\
    \Delta Y_t &= f_Y(N_t) - M_Y Y_t\\
    &\dots
  \end{split}
\end{align}
where $N_t = (X_t,\ Y_t,\ \dots) \in \mathbb{R}^{n}_+$ is a vector of populations, $\Delta X_t := X_{t+1} - X_t$, $\Delta Y_t := Y_{t+1} - Y_t$, and where $f_i:\mathbb{R}^{n}\to\mathbb{R}$ are arbitrary functions.
Here, $M_i=M_i(N_t)\in\mathbb{R}_+$ is a state-dependent fish mortality arising from harvesting (sometimes referred to as *fishing effort*).
Thus, the term $M_X X_t$ is the total $X$ harvest at time $t$.
This formulation of stock recruitment as a discrete finite difference process is common among fisheries, as opposed to continuous time formulations which involve instantaneous growth rates.
This growth rate simplifies e.g. the possibly seasonal nature of reproduction (which would need to be accounted for in a realistic continuous-time model) by simply considering the total recruitment experienced by the population over a full year.

@@

```{r}
#| label: may
#| fig.cap: "The fixed point diagram for the model associated to $f_{\\text{May}}$ as a function of varying the parameter $\\beta H$.
#| That is, these are the zeroes of the function $f_{\\text{May}}$. 
#| Stable fixed points (also known as attractors) are plotted using a solid line, while the unstable fixed point is shown as a dotted line."
#| out.width: "6.5in"

@@

Consider the following three species recruitment model which generalizes eq. \eqref{eq:may}:
\begin{align}
\label{eq:3d model}
\begin{split}
  \Delta X_t &= L(X_t;\ r_X, K_X) - F(X_t, Z_t;\ \beta, c) - c_{XY}X_tY_t,\\
  \Delta Y_t &= L(Y_t;\ r_Y, K_Y) - D F(Y_t, Z_t;\ \beta, c) - c_{XY}X_tY_t,\\
  \Delta Z_t &= \left(f(X_t + DY_t\right) - d_Z)Z_t.
\end{split}
\end{align}
It contains three populations, the state at time $t$ is $(X_t, Y_t, Z_t)$. 
Species $Z$ preys on both $X$ and $Y$, while the latter two compete for resources.
There are ten parameters in this model: 
The growth rate and carrying capacity, $r_X$, $K_X$, $r_Y$ and $K_Y$, of $X$ and $Y$.
A parameter $c_{XY}$ mediating a Lotka-Volterra competition between $X$ and $Y$.
A maximum predation rate $\beta$ and a predation rate half-maximum biomass $c$ specifying how $Z$ forages on $X$ and $Y$.
A parameter $D$ regulating a relative preference of $Z$ to prey on $Y$.
Finally, a death rate $d_Z$ and a parameter $f$ scaling the birth rate of $Z$.
